---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bandsfdp

<!-- badges: start -->
<!-- badges: end -->

This package provides several different upper prediction bands on the FDP in the list of discoveries resulting from competition-based setups (see [Ebadi et al. (2022)](https://arxiv.org/abs/2302.11837)). Such setups include target-decoy competition (TDC) in computational mass spectrometry and the knockoff construction in regression.

The band provides an upper prediction bound on the FDP: given a rejection threshold returned by TDC, a list of target/decoy labels, and a desired confidence level $1 - \gamma$, these functions return a real number $\eta$ such that the FDP in the list of discoveries is $\leq \eta$ with probability $\geq 1 - \gamma$.

## Installation

You can install the development version of bandsfdp from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("uni-Arya/bandsfdp")
```

## Usage

The following inputs, pertaining to the competition-based procedure, are required:

1. One, or several, rejection `thresholds` (the user may choose their own threshold(s) as well).
2. A list of `labels` ($-1$ for a decoy win, $1$ for a target win) that are ordered so the corresponding winning scores are decreasing.

Note to compute the standardized band (`stband()`) and uniform band (`uniband()`), the user must download a set of pre-computed Monte-Carlo data tables using `devtools::install_github("uni-Arya/fdpbandsdata")`.

### With a Single Decoy Score (as in TDC)
 
By default, each band assumes the use of a single decoy score, as in TDC. In this case, in addition to the `thresholds` and `labels`, one must specify the confidence parameter `gamma` (for a `1 - gamma` confidence level). Note that the functions currently support `gamma = 0.01, 0.025, 0.5, 0.1, 0.8, 0.5`, but more data can be generated using the source code available at [GitHub](https://github.com/uni-Arya/fdpbandsdata). For the standardized and uniform band, the value of `alpha` (the FDR tolerance) must also be specified.

```{r bands-single}
suppressPackageStartupMessages(library(bandsfdp))
set.seed(123)

if (requireNamespace("fdpbandsdata", quietly = TRUE)) {
  thresholds <- c(250, 500, 750, 1000)
  labels <- c(
    rep(1, 250),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.9, 0.1)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.5, 0.5)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.1, 0.9))
  )
  alpha <- 0.05
  gamma <- 0.05
  
  print(stband(thresholds, labels, alpha, gamma))
  print(uniband(thresholds, labels, alpha, gamma))
  print(krband(thresholds, labels, gamma))
}
```

### With Multiple Decoy Scores

For competition-based setups that utilizes multiple decoys, one must specify two additional parameters $c \leq \lambda$ of the form $k/(d + 1)$ where $d$ is the number of decoy scores used for each hypothesis and $1 \leq k \leq d$ is an integer. 

For example, if we use $3$ decoy scores for each hypothesis, we may take $c$ and $\lambda$ to be $1/4$, $1/2$, or $3/4$, subject to $c \leq \lambda$. Briefly, the value of $c$ determines the ranks in which the target score is considered "winning." E.g., if $c = 1/4$, $H_i$ is labelled as a target win whenever its corresponding target score is the highest ranked score among all targets/decoys for that hypothesis. Similarly, $\lambda$ determines when the target score is "losing." For more details, see [Ebadi et al. (2022)](https://arxiv.org/abs/2302.11837).

```{r bands-multiple}
suppressPackageStartupMessages(library(bandsfdp))
set.seed(123)

if (requireNamespace("fdpbandsdata", quietly = TRUE)) {
  thresholds <- c(250, 500, 750, 1000)
  labels <- c(
    rep(1, 250),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.9, 0.1)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.5, 0.5)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.1, 0.9))
  )
  alpha <- 0.05
  gamma <- 0.05
  c <- 0.25
  lambda <- 0.25
  
  print(stband(thresholds, labels, alpha, gamma, c, lambda))
  print(uniband(thresholds, labels, alpha, gamma, c, lambda))
  print(krband(thresholds, labels, gamma, c, lambda))
}
```

### Interpolated Bands

By default, all bands are interpolated, which requires the computation of a running maximum. This generally results in a slightly tighter band, but at the cost of computational power. We recommend the use of `interpolate = TRUE`, unless it is too time-consuming. The code below shows an example using non-interpolated bands.

```{r bands-interpolation}
suppressPackageStartupMessages(library(bandsfdp))
set.seed(123)

if (requireNamespace("fdpbandsdata", quietly = TRUE)) {
  thresholds <- c(250, 500, 750, 1000)
  labels <- c(
    rep(1, 250),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.9, 0.1)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.5, 0.5)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.1, 0.9))
  )
  alpha <- 0.05
  gamma <- 0.05
  c <- 0.25
  lambda <- 0.25
  
  print(stband(thresholds, labels, alpha, gamma, c, lambda, interpolate = FALSE))
  print(uniband(thresholds, labels, alpha, gamma, c, lambda, interpolate = FALSE))
  print(krband(thresholds, labels, gamma, c, lambda, interpolate = FALSE))
}
```

### Simultaneous FDP bounds

Rather than compute a bound on the FDP in TDC's list of discoveries, one may be interested in computing a bound on the FDP on the top $i$ hypotheses for all $i = 1, \ldots, n =: [n]$, where $n$ is the number of hypotheses (or for $i$ in a subset of $[n]$). In this case, the function `simband()` may be used. This requires the following arguments:

- The set of (ordered) `labels`, confidence parameter `gamma`, parameters `c` and `lambda`, and a boolean `interpolate`, as described in the previous sections.
- A character argument `type` which is either `"stband"` or `"uniband"`, specifying the type of band to be used to compute simultaneous FDP bounds.
- The set of desired `indices` (denoted by $i$ above).
- The maximum number of decoy wins considered for the bands `d_max` (defaults to `NULL`, in which case is automatically computed using `max_fdp` below).
- The maximum considered FDP for the simultaneous bands `max_fdp` (defaults to `max_fdp = 0.5`).

The arguments `d_max` and `max_fdp` control the rate at which the simultaneous bounds are increasing. We refer the reader to Section 3 of [Ebadi et al. (2022)](https://arxiv.org/abs/2302.11837) for more details. Note that `max_fdp` is the $\alpha$ used to calculate $d_\infty$.

```{r bands-simultaneous}
suppressPackageStartupMessages(library(bandsfdp))
set.seed(123)

if (requireNamespace("fdpbandsdata", quietly = TRUE)) {
  set.seed(123)
  labels <- c(
    rep(1, 250),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.9, 0.1)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.5, 0.5)),
    sample(c(1, -1), size = 250, replace = TRUE, prob = c(0.1, 0.9))
  )
  gamma <- 0.05
  print(simband(labels, gamma, "stband", indices = c(100, 250, 500, 1000)))
}
```

